---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
---

This will contain the future R coding exercise.

# **Loading packages and calling data**
```{r}
#Load Packages: dlslabs and tidyverse 
library("dslabs")
library("tidyverse")
#Help for gapminder data
help(gapminder)
#Overview data structure
str(gapminder)
#Get data summary
summary(gapminder)
#Determine gapminder type of object
class(gapminder)
```

# **Processing Data (Target Variable)**
```{r}
#Create a new object "africadata" in order to analyzed African countries
africadata <- gapminder %>% filter(continent=="Africa")
#Overview data structure
str(africadata)
#Get data summary
summary(africadata)

#Create new objects containing infant_mortality and life_expectancy
mortality_lifexp <- africadata %>% select(infant_mortality, life_expectancy)

#Create new objects containing population and life_expectancy
population_lifexp <- africadata %>% select(population , life_expectancy)

#Overview data structure
str(mortality_lifexp)
str(population_lifexp)

#Get data summary
summary(mortality_lifexp)
summary(population_lifexp)
```

# **Visualization: Plot creation of the new objects/variables**
```{r}
#Plot life_expectacy vs Infant_mortality
ggplot(mortality_lifexp, aes(infant_mortality, life_expectancy)) + geom_point()
#This plot shows a negative correlation between infant mortality and life expectancy. 226 rows were removed due to missing values

#Plot Life_expectancy Vs Population(log scale)
ggplot(population_lifexp, aes(population, life_expectancy)) + geom_point() + scale_x_log10()
#This plot shows a positive correlation between Population and Life expectancy. According with africadata, the frame time when the data was collected was between 1960 to 2016, during this time population size and life expectacy got increse. This is the reason of the streaks. ** 51 rows were removed from the plot due to missing values.
```
# **Cleaning Data**
```{r}
#Finding the year have missing data for infant mortality
africadata_na <- africadata %>% select(year, infant_mortality) %>% filter(is.na(infant_mortality))
africadata_na
#1960 - 1981 and 2016 have missing data.

#To avoid missing data a new variable will be created only with data from year 2000
africadata_2000 <- africadata %>% filter(year==2000)

#Overview data structure
str(africadata_2000)

#Get data summary
summary(africadata_2000)
```
# **Visualization: Plotting clean data**
```{r}
#Plots using only data from year 2000
#Plots will be showed the variable correlation for year 2000
#Plot life expectancy vs infant mortality 
ggplot(africadata_2000, aes(infant_mortality, life_expectancy)) + geom_point()
#The plot shows negative correlation between infant mortality and life expectancy.

#Plot Life_expectancy Vs Population(log scale)
ggplot(africadata_2000, aes(population, life_expectancy)) + geom_point() + scale_x_log10()
#The  plot does not shows a specific relationship between the population size and the life expectancy.
```
# **Fitting linear regression**
```{r}
#We will use lm  function for fit a lineal model with infant mortality as a predictor
fit1<- lm(life_expectancy ~ infant_mortality, data=africadata_2000)

#We will use lm  function for fit a lineal model with population as a predictor
fit2<- lm(life_expectancy ~ population, data=africadata_2000)

#Get the summary
summary(fit1)
summary(fit2)

#Based on the outcomes of the two models and using their p-values, we can conclude that the infant mortality as a predictor (p-value<0.05) shows a significant correlation with life expectancy. The model with population as a predictor has a p-value>0.05, suggesting that is not a good predictor of life expectancy.

```

### Part 2 - Data Manipulation

Here we are going to look at population and GDP in India to see if they are good indicators of life expectancy.

```{r}
library(broom)

#This filters the data so it only holds data from India
Indiadata <- filter(gapminder, country=="India")

```

Next we create two data frames containing the variables we are looking at.

```{r}
IndiaPOPLE <- select(Indiadata, population, life_expectancy)

IndiaGDPLE <- select(Indiadata, gdp, life_expectancy)

summary(IndiaPOPLE)
summary(IndiaGDPLE)

```

### Part 2 - Plotting

```{r}
ggplot(IndiaPOPLE, aes(x = population, y = life_expectancy)) +
      geom_point() +
      xlab("Population") +
      ylab("Life Expectancy")

ggplot(IndiaGDPLE, aes(x = gdp, y = life_expectancy)) +
      geom_point() +
      xlab("GDP") +
      ylab("Life Expectancy")

```

Both of these have positive associations, the first graph with population and life expectancy in particular has a more stable increase over time, whereas the GDP graph has a more exponential increase.

### Part 2 - Linear Models


```{r}
#These create the linear models with life expectancy as the outcome variable and population/gdp as the predictor

fit3 <- lm(population ~ life_expectancy, IndiaPOPLE)

fit4 <- lm(gdp ~ life_expectancy, IndiaGDPLE)



summary(fit3)
summary(fit4)


```

Both have p-values<0.05, so they are both statistically significant predictors of life expectancy.

```{r}
tidy(fit3)
tidy(fit4)
```