---
title: "Tidy Tuesday"
output: 
  html_document:
    toc: FALSE
---
library(tidyverse)
library(here)
library(dplyr)

This assigment is focus is work with dataset from Tidy Tuesday (09/28/2021)

# Load the data - Week 40 09/28/21
```{r}
papers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')

authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')

programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')

paper_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')

paper_programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')
```
## Understanding the data
```{r}
glimpse(papers)
glimpse(authors)
glimpse(programs)
glimpse(paper_authors)
glimpse(paper_programs)

head(programs)
head(papers)
head(paper_programs)

paper_programs%>%
  count(program)
```
The main goal is analyze the category of National Bureau of Economic Research working papers over the time.

```{r}
papers_joined <- paper_programs %>%
  left_join(programs) %>%
  left_join(papers) %>%
  filter(!is.na(program_category)) %>%
  distinct(paper, program, program_category, year, title)

var1<-papers_joined %>% 
  count(program_category)
var1


```
Plot the results
```{r}
P1<- ggplot(data=papers_joined) +
  geom_bar(
    mapping = aes(x=program_category , fill=program_category),
    show.legend = FALSE,
    width = 1)+
    theme(aspect.ratio = 1)+
    labs(x= NULL, y= NULL)
P1 + coord_flip()
P1 + coord_polar()
```
Now that we noticed that the program category behavior, we can check amount of papers through the years.
```{r}
var2<- papers_joined %>%
      subset(select= c(year, program_category)) %>%
      group_by(year)%>%
      count(program_category) %>%
      group_by(year) %>% arrange(year)

#Plot Program category over the time 
P2<- ggplot(data= var2, mapping= aes(x=year, y=n, color= program_category)) + 
  geom_point() +
  geom_smooth()+
  xlab("Year") +
  ylab("Number of Articles")
P2    
```
From the graph we can see that the category Micro has been have a huge increse in the last 20 years, could be interesting to analyze what are the programs with more influence inside of Micro category.
```{r}

head(papers_joined)
var3<- filter(papers_joined, program_category=="Micro")


P3<- ggplot(data=var3) +
  geom_bar(
    mapping = aes(x=program , fill=program),
    show.legend = FALSE,
    width = 1)+
    theme(aspect.ratio = 1)+
    labs(x="Program", y= "Number of Papers")
P3 + coord_flip()
```
The graph shows that Public Economic and Labor Studies are the programs with major number of papers inside of Micro category.